---
title: "DATA 621 Final Project"
author: "Critical Thinking Group 1"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document:
    toc: yes
    toc_depth: 4
  html_document:
    toc: yes
    toc_depth: '2'
    smart: no
  word_document:
    toc: yes
    toc_depth: '2'
---
\pagebreak
\begin{center}
\bigskip
\bigskip
\bigskip
Prepared for:\\
\medskip
Prof. Dr. Nasrin Khansari\\
\smallskip
City University of New York, School of Professional Studies - Data 621\\
\bigskip
DATA 621 – Business Analytics and Data Mining\\
\medskip
Home Work 5\\
\medskip
Prepared by:\\
\medskip
Critical Thinking Group  1\\ 
\medskip
Vic Chan\\ 
\smallskip
Gehad Gad\\
\smallskip
Evan McLaughlin\\  
\smallskip
Bruno de Melo\\
\smallskip
Anjal Hussan\\
\smallskip
Zhouxin Shi\\
\smallskip
Sie Siong Wong\\
\end{center}

\pagebreak

```{r message=FALSE, warning=FALSE, echo=FALSE}
if(!require('tidyverse')) (install.packages('tidyverse'))
```


\newpage

# Introduction
For the final project we will be using different predictive modeling techniques that we have learned throughout the course and using real world data in order to predict if a person has heart disease based on certain factors. The dataset for this final project [Heart Failure Prediction Dataset](https://www.kaggle.com/fedesoriano/heart-failure-prediction) can be found on Kaggle. Each observation in this dataset represents a person's health history, including their age, sex, cholesterol levels, etc. The dataset includes a total of 918 distinct individuals gathered from different countries and agencies. This dataset includes 12 different features, of categorical and/or continous values, of an individual health record, including if the individual has heart disease or not.

## Background
Cardiovascular diseases is the number 1 cause of death globally. The WHO records show that cardiovascular diseases accounted for 31% of all deaths worldwide in 2016.  According to the CDC, nearly 6.2 million adults in the United States suffer heart failure, and in 2018 alone, heart failure was mentioned on 379,800 death certificates. Also, the treatment cost of health care services and medicines is costly and estimated about 31 billion in 2012. Early detection and management for heart disease could be effective in reducing the incidence of heart failure. 

Note that heart failure occurs when the heart cannot pump enough blood and oxygen to support other organs in the body. With this dataset, we would like to see if we can develop a good model to predict if a person has heart disease and what factors can be attributed to heart disease most directly. We will be tackling this question with the usage of different regression techniques and algorithms learned from this class.

# Data Exploration
```{r}
# TODO: Split the dataset into a train and evaluation dataset
dataset = read.csv('https://raw.githubusercontent.com/ahussan/DATA_621_Group1/main/Final/heart.csv')

train_data = dataset
eval_data = dataset
```

# Data Preperation
In this section we will be looking at different ways to prepare the data for modeling and will be showing the different steps and reasons.

## Missing Data
This dataset seems to not have any missing data which is quite rare for a dataset. This means that there is not going to be a need to remove the NA rows or to replace the NA items with the median or mean.
```{r}
na_count = sapply(train_data, function(y) sum(is.na(y)))
na_count = data.frame(na_count)
na_count
```

## Duplicate Row
We will be wanting to remove all duplicate rows in the data set in order to make sure that we will not be skewing results when creating the model
```{r}
train_data = train_data %>% distinct()
```

It seems like that our dataset does not have any duplicates as we have the same amount of rows that we started with.

## Distribution of Heart Disease
Next we will be looking at how many data points do we have where the client has or does not have heart disease. We would like to make sure that the dataset is has a balance of people with and without heart disease in order to create an accurate model.

```{r}
train_data %>%
  count(HeartDisease)
```

As we can see from the result the dataset is also balanced as we have 410 datapoints which dont have heart disease and 508 data points with heart disease.

## Outliers



# Model Creation

# Model Selection

# Conclusion'
